### 2022-02-03 paper.tex
first commit

\documentclass{article}
\usepackage[backend=biber, annotation]{biblatex-chicago}
\usepackage[margin=1in]{geometry}
\addbibresource{paper.bib}

\begin{document}

  \section{Topic that I'm currently considering}

  The current topic that I'm considering is Pakistani music. Although currently
  I'm not very knowledgeable about the subject, I do have a vague idea of what I
  want to study since I am from the region (Pakistan), have a basic
  understanding of its history and also some understanding of the language.
  Currently, I'm thinking of taking a few approaches, Firstly, there is the
  colonial aspect. Since Pakistan was ruled by Britain, I'm considering finding
  music that helps to show how colonialism affected pakistan or how the people
  of the region coped with it. The second aspect is the religious one. In Sufi
  Islam, there are mystical pieces of music called the Qawwalis which are
  somewhat popular. Trying to understand how that was a tool used to cope with
  colonialism could also be illuminating.

  \nocite{*}
  \printbibliography
\end{document}

### 2022-04-18 paper.tex
2 more annotations
introduction and background

@@
 \documentclass{article}
-\usepackage[backend=biber, annotation]{biblatex-chicago}
+\usepackage[authordate, backend=biber, annotation]{biblatex-chicago}
 \usepackage[margin=1in]{geometry}
+\usepackage[]{epigraph}
 \addbibresource{paper.bib}
 
+\usepackage [english]{babel}
+\usepackage [autostyle, english = american]{csquotes}
+\MakeOuterQuote{"}
+
+\title{An Oral History Project: Coke Studio and the Revitalization of Eastern Music}
+\date{April 2022}
+\author{Sarim Aleem}
+
+\setlength\epigraphwidth{.8\textwidth}
+\setlength\epigraphrule{0pt}
+
   \begin{document}
+  \maketitle
 
-  \section{Topic that I'm currently considering}
+  \section{Introduction}
 
-  The current topic that I'm considering is Pakistani music. Although currently
-  I'm not very knowledgeable about the subject, I do have a vague idea of what I
-  want to study since I am from the region (Pakistan), have a basic
-  understanding of its history and also some understanding of the language.
-  Currently, I'm thinking of taking a few approaches, Firstly, there is the
-  colonial aspect. Since Pakistan was ruled by Britain, I'm considering finding
-  music that helps to show how colonialism affected pakistan or how the people
-  of the region coped with it. The second aspect is the religious one. In Sufi
-  Islam, there are mystical pieces of music called the Qawwalis which are
-  somewhat popular. Trying to understand how that was a tool used to cope with
-  colonialism could also be illuminating.
+  \epigraph{\itshape One day, I suddenly discovered eastern music. It was a very
+  strange day for me, you know when something inside just goes click.}{--- Rohail Hyatt}
+
+  Like Rohail Hyatt, most Pakistani's aren't aware of the one thousand year
+  heritage of music and culture that in the diaspora. Most would ignore the rich
+  classical music scene for being boring or inaccessible, and would most likely
+  rather opt for bollywood style dance videos, a dime a dozen in the country.
+
+  But as Hyatt, the leader of a reignition of a rich and vibrant traditional
+  culture would discover, there is an incredible philosophical and poetic depth
+  to traditional music.  But little did he expect that the reignition of that
+  lost culture would not come from the local film or entertainment industry.
+  Rather, it would come from a well known Western soft drink company, initially
+  trying to outcompete its famed rival through a small marketing stint: Coca
+  Cola. And, he would spearhead it.
+
+  The introduction of Coke Studio in Pakistan has led to the revitalization of
+  traditional culture by putting a reintroducing endangered languages, putting a
+  spotlight on traditional Eastern instruments, and bringing traditional Sufi
+  culture into the modern day.
+
+  \section{Background}
+
+  It began as a marketing project in Pakistan. Pepsi, Coca Cola's main rival
+  over 70\% of the soft drink market share. Pepsi's success was largely due to
+  its association with Cricket and Music in the mind of most Pakistanis: "Be it
+  through sponsorships, endorsements with big cricket stars, investing in new
+  emerging cricketing and music talent, Pepsi was successful in becoming a
+  global brand with local connection" \autocite{Chibmedium}. With growth 
+  potential low in more developed countries, Coca Cola saw an opportunity to 
+  grow its product in developing countries.
+
+  Still though, most Pakistani were not terribly interested in traditional folk 
+  music given the wave of westernization with the internet. But Rohail Hyatt had a 
+  different idea. Strike at Pakistani cultural pride in the face of globalization, but 
+  give audiences a taste of modernity within the folk music to draw them in. Thus, 
+  \textit{Fusion} was born, a term coined by Hyatt to refer to the mix between modern pop music
+  and traditional folk \autocite{dhanwani2014coke}.
+
+  Thus, Coke studio was born; blending East and West when it came to musical instruments, culture,
+  and musical character. This blend would not only revitalize traditional culture but innovate on it.
 

### 2022-04-20 paper.tex
partial progress on sufism

@@
 
-  Thus, Coke studio was born; blending East and West when it came to musical instruments, culture,
-  and musical character. This blend would not only revitalize traditional culture but innovate on it.
+  Thus, Coke studio was born; blending East and West when it came to musical
+  instruments, culture, and musical character. This blend would not only
+  revitalize traditional culture but innovate on it.
+
+  \section{A Reintroduction of Sufi Culture}
+
+  There are three main branches in the religion of Islam. Sunni - the normative
+  version of Islam practiced by the majority of the world. Shi'a - a smaller but
+  not insignificant sect with minor differences in religious interpretation. And
+  finally, there is Sufi Islam. Although technically Sufi Islam is not
+  considered by some to be a separate branch of Islam, many of their practices
+  are not found in normative Islam and thus, for the purposes of this paper they
+  will be considered a different "branch" or "sect".
+
+  Sufi Islam is commonly defined "Islamic Mysticism" and tends to emphasize the 
+  spiritual aspects of the religion more so than the legalistic, or worldy
+  elemnts. For most orientalists, the religion is thus seen as the more 
+  moderate, "peaceful" interpretation of islam \autocite{beg2020fizzy}. This of
+  course incorrectly implies that traditional mainstream islam is oppressive,
+  but nevertheless the perception of Westerners ironically have a large
+  influence on what local Pakistanis wish to express their culture. As a
+  consequence, embracing the Sufi genre of music - the Qawwali - is a nice
+  middle ground in appeasing to a moderate image of Islam while simultaneously 
+  appealing to the more religiously inclined in the nation who are skeptical of
+  the value of music in society \autocite{beg2020fizzy}.
+
+  The Qawwali is a musical manifestation of Sufism. It's a piece of religious music 
+  that tends to be 10-20 minutes long, but can range up to an hour and is typically 
+  based on religious poetry. The piece starts out slowly, with a couple words chanted
+  in the begining, but quickly accelerates to a climax.
 

### 2022-04-22 paper.tex
somewhat finished rough draft (actual garbage tho)

@@
 
-  \nocite{*}
+  Coke studio's rendition of the 800 year old famous poem "Chaap Tilak" is a
+  quintissential example of such Qawwali \autocite{Chaaptilaak}. The song starts
+  slow, and the singers are chanting one word over the course of 10-15 seconds.
+  The slow introduction serves the purpose to "prime" those who don't understand
+  the language, as the music can be appreciated by anyone. The song slowly
+  accelerates to a climax so intense the singers start sweating profusely.
+
+  What's important to note here is not only the song itself, but its reception. 
+  The video itself has almost \textit{fifty million views} on Youtube, extremely
+  high for a rendition of a classical poem. Furthermore, the comment section
+  seems to affirm the sublime superiority of classical culture over modernity.
+  One user writes: "Maturity is when you understand classical music is apart
+  from all genre." Another comments points to the fact that the music trancends political
+  differences in the subcontinent. One user writes: "I'm a Pakistani. but I can
+  see here most of the comments are from Indians It shows Indian are really
+  music lover and they know how to appreciate talent."
+
+  Part of the reason that the genre of Sufi music became so popular was that it
+  largely touched on the spiritual aspects of religion. Another part is that the
+  poem predates the beginning of political differences in the Indian
+  subcontinent. And yet another is the fact that The video is freely available
+  on youtube, arguably one of the largest if not the largest video streaming
+  site in the world. Combined with Coca Cola's large marketing budget, it's not then 
+  surprising the song became the hit it did.
+
+  \section{Women and Young Performers in Coke Studio}
+
+  Although Coke Studio has bought Sufi music into the mainstream, there are aspects
+  of it that differ greatly from its traditional performance.
+
+  On a reexamination of the video above, one may notice that Abida Parveen is a woman. 
+  While it's not particularly unusual for a woman to sing in a Western context, it's 
+  certainly the case that when it comes to Islamic religious veneration, women are 
+  traditionally not "leaders." For this reason, men traditionally led Qawwali's.
+
+  Take for instance, Naz Warsi, who is a trained female musician in the Sufi tradition.
+
+  \begin{quote}
+
+    Warsi told the reporter that although her desire - similar to other female
+    performers - is to perform qawwali, they are not granted permission.
+    Although Warsi has obtained training in Sufi classical singing and has
+    earned her name in the genre, it is twice the struggle for her to take her
+    place in the world of qawwali. She is not only a woman but a woman who does
+    not possess the “proper” musical lineage (BBC Radio Urdu, n.d.).  It said
+    that only men should be permitted to sing qawwali but that men who have
+    inherited the profession should be the only ones to perform. The role of
+    females is actively discouraged within sacred dargah culture. Only male
+    heirs are to succeed their fathers, making paternal lineage the trait that
+    allows a singer to become a performer in a Sufi dargah
+    \autocite{beg2020fizzy}.
+
+  \end{quote}
+
+  In order to appeal the a large segment of the modern population, Coke Studio
+  has taken to the approach of of intermixing traditional Sufi music with modern
+  styles. One way it does this is by not only adding women, but also young
+  people not traditionally thought to have sung the songs.
+
+  Take for example, arguably the most popular Coke Studio hit: "Afreen."
+
+  Afreen is a quintissential example of Coke studio subverting expectations when
+  it came to artist selection. The song was originally sung by the dubbed "King
+  of Qawwali" Nusrat Fateh Ali Khan. The cover on the other hand was not only
+  sung by his nephew, the esteemed Rahat Fateh Ali Khan, but also by Momena
+  Mostehsan: a young woman raised in the United States. The combining of
+  cultures brings Sufi culture into the modern era, and thus there's a sense of
+  innovation in the music. Furthermore, not only is the artist selection crucial,
+  but also their appearance.
+
+\begin{quote}
+  what garnered more attention sometimes were the looks of the artist, than the
+  artist himself/ herself. One would often find the folk musician accompanied by a
+  female singer in jeans, cropped leather jacket, painted nails, which was indeed
+  a counter narrative to the burkha wearing Islam women (\autocite{Chibmedium}).
+\end{quote}
+
+Thus, although Coke Studio brings old Sufi culture into the modern day, it is not done
+without changing who performs the songs.
+
+\section{Homogenization of Language}
+
+Although Coke Studio has bought an incredible amount of innovation of into the
+music scene, there are reminders that at the end of the day the studio is just a
+marketing stint for a multi-national coorporation. As such, there can be serious
+oversights when it comes to what's performed and what isn't. One such oversight is 
+clearly present when it comes to the issue of language.
+
+Unlike the United States, which is mostly homogenous in terms of linguistic
+diversity, the hindustani region has an incredible amount of languages and
+subcultures splintered in.  There are over 30 languages alone with over 1
+million speakers in the subcontinent, and about 122 languages with more than
+10,000 speakers.
+
+It's then quite unfortunate that Coke Studio is lacking in the language
+department, and fails to cover the linguistic diversity of the subcontinent.
+
+Rodrigo Chocano, a Peruvian enthnomusicologist writes:
+
+\begin{quote}
+  Between the two [Urdu and Punjabi], they account for way over two thirds of the whole pie.
+
+  Fifteen languages split the remaining spots. In comparison to Punjabi, the other
+  three major regional languages pale: while Pashto, Balochi and Sindhi together
+  account for over one third of native speakers in Pakistan, they only amount to
+  four percent of the linguistic presence in Coke Studio Pakistan.
+\end{quote}
+
+It's hard to forget that at the end of the day, Coca Cola is a business. It's
+simply more convenient to appeal to the "prestige" languages that are used in
+Poetry (Persian) or Pop Culture (Urdu, Punjabi) than go out of the way to produce 
+music in other languages.
+
+Although this may look harmless at the outset, if Coke Studio is viewed, as
+perhaps many do, as being an musical representation of the sould of the nation,
+then many niche but equally important subcultures can be overlooked, and the
+perception of culture is then formed by a media company that is, for lack of
+better terms, "in it for the money." Unfortunately, money isn't concerned with
+diversity or representation if said diversity doesn't help the bottom line.
+
+But still, the issue becomes worse when poetry from other languages is sung
+incorrectly:
+
+\begin{quote}
+
+ It's not just that certain nuances are getting lost in what these artists are
+ singing. They are actually singing some completely meaningless lines and you
+ can imagine what that means for the transmission of meaning for classical
+ poetry that has been traditionally passed on from one generation to another.
+
+\end{quote}
+
+Arguably the largest music producer in the country is doing a disservice to 
+classic poetry and culture. There is an element of neglect in the rendition of
+popular pieces of classic poetry, perhaps hinting at the fact the the priorities of 
+Coke Studio are slightly superficial.
+
+\section{Conclusion}
+
+Although the Coke Studio discussed, the one popularly known, is the Pakistani
+one, Coca Cola has in fact attempted to establish studios in other countries,
+including India. However, the Indian Coke studio never reached the fame the
+Pakistani did, since it couldn't escape it's consumerist, capitalist, Bollywood
+roots. But, the success of Pakistani Coke Studio points to the fact that in the
+face of the globalization and westernization of Pakistan, there is a sense of
+pride in sublime of classic poetry and music, however imperfect the modern
+rendition of that may be.
+
 \printbibliography

### 2022-05-09 paper.tex
just final proofread and I'm done

@@
 \date{April 2022}
-\author{Sarim Aleem}
+\author{Sarim Aleem \\ Word Count: 2135}
 
@@
   classical music scene for being boring or inaccessible, and would most likely
-  rather opt for bollywood style dance videos, a dime a dozen in the country.
+  rather opt for bollywood style dance videos - a dime a dozen in the country.
 
@@
   trying to outcompete its famed rival through a small marketing stint: Coca
-  Cola. And, he would spearhead it.
+  Cola.
 
-  The introduction of Coke Studio in Pakistan has led to the revitalization of
-  traditional culture by putting a reintroducing endangered languages, putting a
-  spotlight on traditional Eastern instruments, and bringing traditional Sufi
-  culture into the modern day.
+  The introduction of Coke Studio in Pakistan has led to a revitalization of
+  classical music by reintroducing Sufi culture, putting a diverse catalog of
+  artists in music, and having a nontrivial amount of linguistic diversity in
+  songs.  Nevertheless, the production studio has it's flaws characteristic of a
+  large capitalistic coorporation, and oftentimes commoditizes classic culture
+  and neglects it's heritage.
 
@@
 
-  Thus, Coke studio was born; blending East and West when it came to musical
+  And then, Coke studio was born; blending East and West when it came to musical
   instruments, culture, and musical character. This blend would not only
-  revitalize traditional culture but innovate on it.
+  revitalize traditional culture but for better of worse, innovate on it.
 
@@
   are not found in normative Islam and thus, for the purposes of this paper they
-  will be considered a different "branch" or "sect".
+  will be considered a different "branch" or "sect" \autocite{sufibackground}.
 
@@
   elemnts. For most orientalists, the religion is thus seen as the more 
-  moderate, "peaceful" interpretation of islam \autocite{beg2020fizzy}. This of
+  moderate, "peaceful" interpretation of islam. This of
   course incorrectly implies that traditional mainstream islam is oppressive,
   but nevertheless the perception of Westerners ironically have a large
-  influence on what local Pakistanis wish to express their culture. As a
+  influence on what local Pakistanis wish to express their culture \autocite{beg2020fizzy}. As a
   consequence, embracing the Sufi genre of music - the Qawwali - is a nice
@@
   appealing to the more religiously inclined in the nation who are skeptical of
-  the value of music in society \autocite{beg2020fizzy}.
+  the value of music in society \autocite{williams2019soundtrack}.
 
@@
 
+  As with most Sufi poetry, the lyrical content of Chaap Tilak is very divorced from 
+  political reality, and is far more focused on abstract and "higher" concepts:
+
+  \begin{quote}
+
+    The Game of Love
+    The Game of Love
+    If I play with my dear one
+
+    If I win, My sweetheart is Mine
+    If I lose, I'm still with my Dear one
+
+    \dots
+
+    The night of the Prophet's celebration has dawned
+    The night has dawned
+    The night has dawned
+
+    \dots
+
+    With one glance your enchanting eyes
+
+  \end{quote}
+
+  Love and devotion to God featured heavily in the song, and much of Sufi
+  Qawwali is described as eschewing wordly and material life - hardly a
+  controversial message. For this reason, Qawwali's are a popular genre to
+  market to worldwide audiences
+
   What's important to note here is not only the song itself, but its reception. 
@@
 
+  That's not to say that everyone is happy with the reintroduction of Sufism in modern culture. 
+  For many, the reintroduction of Sufi culture seems commodotized. As Amin writes:
+
+  \begin{quote}
+    What happened to qawwali in this process? [After it's greater presence in
+    90's] It was unmoored from its deeper spiritual significance and repackaged
+    as a cultural item, all so that it could be more easily commodified for a
+    global audience. \autocite{sellingsufism}
+  \end{quote}
+
+  For many, Coke Studio adopting the Qawwali was not seen as a victory, but
+  rather a capitalist society appropriating a millenia old tradition for profit.
+  In fact, although the Qawwali comes from the Islamic tradition, "Qawwali is
+  rarely marketed as Islamic music because to do so would risk turning away
+  consumers, as Islam purportedly does not share the same peaceful values as
+  Sufism" \autocite{sellingsufism}. So in essence, a company like Coke Studio 
+  takes the Qawwali, and erases it's cultural heritage when convenient to make 
+  profits - Hardly an ode to authenticity.
+
   Part of the reason that the genre of Sufi music became so popular was that it
@@
   site in the world. Combined with Coca Cola's large marketing budget, it's not then 
-  surprising the song became the hit it did.
+  surprising that Qawwalis became the hits they did in Coke Studio.
 
@@
   sung by his nephew, the esteemed Rahat Fateh Ali Khan, but also by Momena
-  Mostehsan: a young woman raised in the United States. The combining of
+  Mostehsan: a young woman raised in the United States \autocite{afreen}. The combining of
   cultures brings Sufi culture into the modern era, and thus there's a sense of
@@
   female singer in jeans, cropped leather jacket, painted nails, which was indeed
-  a counter narrative to the burkha wearing Islam women (\autocite{Chibmedium}).
+  a counter narrative to the burkha wearing Islam women \autocite{Chibmedium}.
 \end{quote}
 
-Thus, although Coke Studio brings old Sufi culture into the modern day, it is not done
-without changing who performs the songs.
+The inclusion of women and nontraditional artists is important because it
+signals a change in how classical music is performed. The music is stripped down
+to it's bare minimum and reconstructed. And, for the most part it has succeeded.
+Songs with these new and modern artists have hundreds of millions of views on
+youtube, and have captivated the imagination of the diaspora.
 
-\section{Homogenization of Language}
+\section{The language of Coke Studio}
 
@@
 Unlike the United States, which is mostly homogenous in terms of linguistic
-diversity, the hindustani region has an incredible amount of languages and
+diversity, the Hindustani region has an incredible amount of languages and
 subcultures splintered in.  There are over 30 languages alone with over 1
@@
   account for over one third of native speakers in Pakistan, they only amount to
-  four percent of the linguistic presence in Coke Studio Pakistan.
+  four percent of the linguistic presence in Coke Studio Pakistan \autocite{soundscape}.
 \end{quote}
 
-It's hard to forget that at the end of the day, Coca Cola is a business. It's
+That's not to say that Coke Studio has no linguistic diversity. The producers
+are humans after all and do make attempts to put in some sort of linguistic
+diversity.  For example the following languages and scripts have been learnt and
+used by the producers: "Seraiki, Sindhi, Pashto, Punjabi, Brahui, Balochi, Braj
+Bhasha, Marwari, Bengali, Urdu, Persian, Arabic and Turkish" \autocite{herald}.
+That is still a great deal more than is present in bollywood media.
+
+Still, It's hard to forget that at the end of the day, Coca Cola is a business. It's
 simply more convenient to appeal to the "prestige" languages that are used in
 Poetry (Persian) or Pop Culture (Urdu, Punjabi) than go out of the way to produce 
-music in other languages.
+music in other languages \autocite{ethnographicInterview}.
 
@@
 \begin{quote}
-
  It's not just that certain nuances are getting lost in what these artists are
@@
  can imagine what that means for the transmission of meaning for classical
- poetry that has been traditionally passed on from one generation to another.
+ poetry that has been traditionally passed on from one generation to another \autocite{herald}.
+\end{quote}
 
-\end{quote}
+We tend to see unfortunately that there is a sense of neglect. One megahit that 
+coke studio produced is a song by the name "Taj  Da Re Haram"
+\autocite{taajdaar}. The song is the most popular song on the youtube channel by
+view count (over 374 million views as of this writing).  However, out of
+carelessness many of the lyrics are wrong to the point of being nonsensical
+\autocite{herald}.
+
+In some part, this failing can be attributed to the fact that education in
+traditional languages like Urdu have lessened in recent years, as English is a
+far more useful language from a capitalistic standpoint. Popular artists don't 
+fact check their songs for accuracy all the time. This is reminiscent of the 
+previous discussion on the commodotization of Qawwali's. After all, if a great 
+deal of the population is unable to fact check the songs themselves, then it
+doesn't hurt the bottomline too much to overlook careful scrutiny that ensures
+the classic renditions are authentic.
 
@@
 
-Although the Coke Studio discussed, the one popularly known, is the Pakistani
+Although the Coke Studio discussed and the one popularly known is the Pakistani
 one, Coca Cola has in fact attempted to establish studios in other countries,
@@
 
+Coke Studios innovation in the field has been nothing short of incredible.
+Classic Poetry from almost a millenia ago are being popularized and being
+listened to hundreds of millions of times.  However, in the usage of classical
+sources, it is imperative to stay faithful to them, which Coke Studio has failed
+to do. After all, for many, Coke Studio has become more than a marketing stint; 
+it has trancended to Pakistan's oral history project.
+
 \printbibliography
